Consensus cell types across samples from brain and CNS tumors

### 2025-02-20 analyses/cell-type-consensus/exploratory-notebooks/04-brain-and-cns-consensus-celltypes.Rmd
re-render brain notebook

```{r}
# list all results files 
results_files <- list.files(results_dir, pattern = "_consensus-cell-types\\.tsv.\\gz$", full.names = TRUE)

# get project ids from file list and assign as names
project_ids <- stringr::str_remove(basename(results_files), "_consensus-cell-types.tsv.gz")
names(results_files) <- project_ids

# grab only brain/cns projects
brain_project_ids <- c("SCPCP000001", "SCPCP000002", "SCPCP000010", "SCPCP000021", "SCPCP000009")
results_files <- results_files[brain_project_ids]

@@

# read in results and prep data frame for plotting 
all_results_df <- results_files |> 
  purrr::imap(summarize_celltypes) |> 
  dplyr::bind_rows(.id = "project_id") |> 
  # remove multiplexed libraries 
  dplyr::filter(!(library_id %in% multiplexed_libs)) |> 
  # join with sample metadata
  dplyr::left_join(all_metadata_df, by = c("project_id", "library_id"))
```

@@

# get all possible cell types 
# there are exactly 26 of them so we will use alphabet with two greys for unknown/all remaining 
unique_celltypes <- all_results_df |> 
  dplyr::filter(percent_cells_annotation > 0.1, # we're only ever going to plot cell types that are greater than 1%
                !consensus_annotation == "Unknown") |> 
  dplyr::pull(consensus_annotation) |> 
  unique() |>
  as.character()

# define colors for all cell types
all_celltype_colors <- c(
  palette.colors(palette = "alphabet"),
  "grey60", # all remaining
  "grey95", # unknown
  "grey60" # non-immune, use the same as all remaining since they are never in the same plot
)
names(all_celltype_colors) <- c(unique_celltypes, "All remaining cell types", "Unknown", "non-immune")

>>>>

```{r}
# list all results files 
brain_project_ids <- c("SCPCP000001", "SCPCP000002", "SCPCP000010", "SCPCP000021", "SCPCP000009")
brain_results_dirs <- file.path(results_dir, brain_project_ids)
results_files <- brain_results_dirs |> 
  list.files(pattern = "_processed_consensus-cell-types\\.tsv.\\gz$", recursive = TRUE, full.names = TRUE)

@@

# read in results and prep data frame for plotting 
all_results_df <- results_files |> 
  purrr::imap(summarize_celltypes) |> 
  dplyr::bind_rows() |> 
  # remove multiplexed libraries 
  dplyr::filter(!(library_id %in% multiplexed_libs)) |> 
  # join with sample metadata
  dplyr::left_join(all_metadata_df, by = c("project_id", "library_id"))
```

@@

# get all possible cell types 
# there are exactly 26 of them so we will use alphabet with two greys for unknown/all remaining 
unique_celltypes <- all_results_df |> 
  # order by frequency otherwise neurons will be light grey 
  dplyr::arrange(desc(total_cells_per_annotation)) |> 
  dplyr::filter(percent_cells_annotation > 0.1, # we're only ever going to plot cell types that are greater than 1%
                !consensus_annotation == "Unknown") |> 
  dplyr::pull(consensus_annotation) |> 
  unique() |>
  as.character()

# define colors for all cell types
all_celltype_colors <- c(
  palette.colors(palette = "alphabet"),
  "black", # extra cell type 
  "grey60", # all remaining
  "grey95", # unknown
  "grey60" # non-immune, use the same as all remaining since they are never in the same plot
)
names(all_celltype_colors) <- c(unique_celltypes, "All remaining cell types", "Unknown", "non-immune")
